---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

Victoria Schwab is an American fantasy and science fiction writer. She debuted her first novel, "The Near Witch" in 2011 and has spent the past 11 years building on this discography. She has published around 39 books, multiple short stories, and has adapted her stories into cinema. Her first TV Show "First Kill", based on her short story by the same name, was released on Netflix in June 2022. As Victoria Schwab's discography increases, new readers of her work need guidance on which books or series to start their Victoria Schwab journey with. This report, which examines data from Seattle Public Library, will look into the popularity of Victoria Schwabs books and series to offer a good starting point to her new audience. 

Specifically, this report analyzes three important questions. The first questions is on which book series of Victoria Schwab's was the most popular, and which month was the most popular for each series? The second question asks which medium is the most popular and how has this changed over the years? Lastly, this report will answer which of Victoria Schwab's books is the most popular in the past year? To answer these questions, the key variables include:

* Library checkouts:
    * The amount of times a book is checked out of the library is used to quantify a particular books popularity. 
* Time
    * What is the trend of checking out Victoria Schwab books or series? In this report, both year and month are used as time frames to inform on the trend of popularity of each book. 
* Books
    * Which books or series hold the highest popularity? Victoria Schwab writes both stand alone novels and book series. This report looks into each book individually as well as books bundled with their series to give new audience members options of which type to begin their journey with. 
* Mediums
    * Which medium is Victoria Schwabs books typically enjoyed using? This variable encompass' physical books, eBooks, audiobooks, and soundisc's. Understanding the medium of engaging with the book will inform a new reader on whick medium they choose to engage the books with themselves. 

### Summary Information
```{r, echo = FALSE, message = FALSE, results ='hide'}
# relevant calculations
library("dplyr")
library("stringr")
Schwab_df <- read.csv("Schwab_Checkouts_by_Title.csv", stringsAsFactors = F)

Schwab_df$Title <- tolower(Schwab_df$Title)
Schwab_df$Title[str_detect(Schwab_df$Title, "vicious")] <- "Vicious"
Schwab_df$Title[str_detect(Schwab_df$Title, "a conjuring of light")] <- "A Conjuring of Light"
Schwab_df$Title[str_detect(Schwab_df$Title, "a gathering of shadows")] <- "A Gathering of Shadows"
Schwab_df$Title[str_detect(Schwab_df$Title, "new beginnings")] <- "New Beginnings"
Schwab_df$Title[str_detect(Schwab_df$Title, "broken ground")] <- "Broken Ground"
Schwab_df$Title[str_detect(Schwab_df$Title, "this savage song")] <- "This Savage Song"
Schwab_df$Title[str_detect(Schwab_df$Title, "the unbound")] <- "The Unbound"
Schwab_df$Title[str_detect(Schwab_df$Title, "a darker shade of magic")] <- "A Darker Shade of Magic"
Schwab_df$Title[str_detect(Schwab_df$Title, "the near witch")] <- "The Near Witch"
Schwab_df$Title[str_detect(Schwab_df$Title, "second chances")] <- "Second Chances"
Schwab_df$Title[str_detect(Schwab_df$Title, "last wishes")] <- "Last Wishes"
Schwab_df$Title[str_detect(Schwab_df$Title, "the archived")] <- "The Archived"
Schwab_df$Title[str_detect(Schwab_df$Title, "bridge of souls")] <- "Bridge of Souls"
Schwab_df$Title[str_detect(Schwab_df$Title, "city of ghosts")] <- "City of Ghosts"
Schwab_df$Title[str_detect(Schwab_df$Title, "conjuro de luz")] <- "A Conjuring of Light"
Schwab_df$Title[str_detect(Schwab_df$Title, "extraordinary")] <- "ExtraOrdinary"
Schwab_df$Title[str_detect(Schwab_df$Title, "gallant")] <- "Gallant"
Schwab_df$Title[str_detect(Schwab_df$Title, "la vida invisible de addie larue")] <- "The Invisible Life of Addie LaRue"
Schwab_df$Title[str_detect(Schwab_df$Title, "the invisible life of addie larue")] <- "The Invisible Life of Addie LaRue"
Schwab_df$Title[str_detect(Schwab_df$Title, "our dark duet")] <- "Our Dark Duet"
Schwab_df$Title[str_detect(Schwab_df$Title, "tunnel of bones")] <- "Tunnel of Bones"
Schwab_df$Title[str_detect(Schwab_df$Title, "vengeful")] <- "Vengeful"

# rise_df is the rise in checkouts per year and total in all years
rise_df <- Schwab_df %>% 
  group_by(CheckoutYear) %>%
  summarise(Checkouts = sum(Checkouts, na.rm = T)) %>%
  mutate(growth = lag(Checkouts, 1)) %>%  # the year before is in growth_pop
  mutate(growth = ((Checkouts - growth) / growth) * 100) %>%
  mutate(total_growth = ((Checkouts[12] - Checkouts[1]) / Checkouts[1] * 100)) %>%
  filter(CheckoutYear != 2023)

rise_total <- rise_df$total_growth[1]
greatest_rise <- max(rise_df$growth, na.rm = T)
pop_year <- rise_df$CheckoutYear[which(rise_df$growth== greatest_rise)]

# total_checkouts is the total amount of checkouts she has across all years

total_checkouts <- sum(Schwab_df$Checkouts, na.rm = T)

# most_used_medium are the mediums that are used the most across every book
most_used_medium <- Schwab_df %>% 
  group_by(MaterialType) %>%
  summarise(Checkouts = sum(Checkouts, na.rm = T)) %>%
  arrange(desc(Checkouts)) %>%
  mutate(MaterialType = tolower(MaterialType))

# Most_popular_book is a list of which books are most popular by checkouts
most_popular_book <- Schwab_df %>% 
  group_by(Title) %>%
  summarise(Checkouts = sum(Checkouts, na.rm = T)) %>%
   arrange(desc(Checkouts))

# Most_popular_series is a list of which series are the most popular by checkouts
most_popular_series <- Schwab_df %>% 
  mutate(Series = Title) %>%
  group_by(Series) %>%
  filter(CheckoutYear == "2022") %>%
  select(Title, Checkouts, CheckoutMonth) %>%
  mutate(Series = gsub("A Conjuring of Light", "Shades of Magic Series", Series)) %>%
  mutate(Series = gsub("A Darker Shade of Magic", "Shades of Magic Series", Series)) %>%
  mutate(Series = gsub("A Gathering of Shadows", "Shades of Magic Series", Series)) %>%
  mutate(Series = gsub("Bridge of Souls", "City of Ghosts Series", Series)) %>%
  mutate(Series = gsub("City of Ghosts", "City of Ghosts Series", Series)) %>%
  mutate(Series = gsub("City of Ghosts Series Series", "City of Ghosts Series", Series)) %>%
  mutate(Series = gsub("Tunnel of Bones", "City of Ghosts Series", Series)) %>%
  mutate(Series = gsub("This Savage Song", "Monsters of Verity Series", Series)) %>%
  mutate(Series = gsub("Our Dark Duet", "Monsters of Verity Series", Series)) %>%
  mutate(Series = gsub("The Archived", "The Archived Series", Series)) %>%
  mutate(Series = gsub("The Unbound", "The Archived Series", Series)) %>%
  mutate(Series = gsub("Vicious", "Villains Series", Series)) %>%
  mutate(Series = gsub("Vengeful", "Villains Series", Series)) %>%
  mutate(Series = gsub("ExtraOrdinary", "Villains Series", Series)) %>%
  mutate(Series = gsub("New Beginnings", "Everyday Angel Series", Series)) %>%
  mutate(Series = gsub("Last Wishes", "Everyday Angel Series", Series)) %>%
  mutate(Series = gsub("Second Chances", "Everyday Angel Series", Series)) %>%
  filter(Series != "Broken Ground", Series != "The Invisible Life of Addie LaRue", Series != "Gallant", Series != "The Near Witch") %>% 
  group_by(Series) %>%
  summarise(Checkouts = sum(Checkouts, na.rm = T))%>%
   arrange(desc(Checkouts))

```

Victoria Schwab's fame as a writer has been steadily increasing throughout the years with a **`r rise_total`%** increase since 2011 when she released her first novel "The Near Witch." The year with the highest marked growth was in **`r pop_year`** with a **`r greatest_rise`%** increase from the year prior. Her most popular book at this point of her career is **`r most_popular_book$Title[1]`** with **`r most_popular_book$Checkouts[1]`** checkouts out of the **`r total_checkouts`** checkouts of Victoria Scwabs work. For a newcomer to Victoria Schwabs discography looking for a stand alone book, **`r most_popular_book$Title[1]`'s** popularity indicates that **`r most_popular_book$Title[1]`** might be a good place to start. Victoria Schwabs second most checked out books is **`r most_popular_book$Title[2]`** with **`r most_popular_book$Checkouts[2]`** checkouts in total. If the first book does not appeal to the reader, **`r most_popular_book$Title[2]`** would likely be another good place to start her discography.

In terms of most popular series, **`r most_popular_series$Series[1]`** is the series with the highest amount of checkouts, with **`r most_popular_series$Checkouts[1]`** checkouts in total. Her second most popular series is the **`r most_popular_series$Series[2]`** with **`r most_popular_series$Checkouts[2]`** checkouts. These two series would be a good starting point for a reader looking for a series over a stand alone novel. 

When it comes to the method of engaging with Victoria Schwabs works, the most used medium for people checking out her work from the library is an **`r most_used_medium$MaterialType[1]`**. Out of the **`r total_checkouts`** checkouts of Victoria Scwabs work, **`r most_used_medium$Checkouts[1]`** of these checkouts have been in **`r most_used_medium$MaterialType[1]`** format. The second most common method for accessing her material is through **`r most_used_medium$MaterialType[2]`**, with **`r most_used_medium$Checkouts[2]`** out of the **`r total_checkouts`** total checkouts of Victoria Schwab books checked out being **`r most_used_medium$MaterialType[2]`'s**. For beginners without a preference in methods for engaging in material, these two methods are seen as the preferred method for accessing her material to library goers and worth considering engaging in her discography with. 

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:
``````{r, echo = FALSE, warning = FALSE, message = FALSE}
# Question: Which month was most popular for each of Victoria Schwab's series in the past full year? Which is Victoria Schwab's most popular book series? 

#load Libraries
library("dplyr")
library("ggplot2")
library("RColorBrewer")

# load Data Frame
Schwab_df <- read.csv("Schwab_Checkouts_by_Title.csv", stringsAsFactors = F)
Schwab_df$Title <- tolower(Schwab_df$Title)
Schwab_df$Title[str_detect(Schwab_df$Title, "vicious")] <- "Villains Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "a conjuring of light")] <- "Shades of Magic Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "a gathering of shadows")] <- "Shades of Magic Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "new beginnings")] <- "Everyday Angel Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "broken ground")] <- "N/A"
Schwab_df$Title[str_detect(Schwab_df$Title, "this savage song")] <- "Monsters of Verity Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "the unbound")] <- "The Archived Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "a darker shade of magic")] <- "Shades of Magic Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "the near witch")] <- "N/A"
Schwab_df$Title[str_detect(Schwab_df$Title, "second chances")] <- "Everyday Angel Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "last wishes")] <- "Everyday Angel Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "the archived")] <- "The Archived Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "bridge of souls")] <- "City of Ghosts Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "city of ghosts")] <- "City of Ghosts Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "conjuro de luz")] <- "Shades of Magic Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "extraordinary")] <- "Villains Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "gallant")] <- "N/A"
Schwab_df$Title[str_detect(Schwab_df$Title, "la vida invisible de addie larue")] <- "N/A"
Schwab_df$Title[str_detect(Schwab_df$Title, "the invisible life of addie larue")] <- "N/A"
Schwab_df$Title[str_detect(Schwab_df$Title, "our dark duet")] <- "Monsters of Verity Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "tunnel of bones")] <- "City of Ghosts Series"
Schwab_df$Title[str_detect(Schwab_df$Title, "vengeful")] <- "Villains Series"

# modify Data Frame
popular_df <- Schwab_df %>%
  filter(CheckoutYear == "2022") %>%
  select(Title, Checkouts, CheckoutMonth) %>%
  mutate(Series = Title) %>%
  group_by(Series, CheckoutMonth) %>%
  summarise(Checkouts = sum(Checkouts))%>%
  filter(Series != "N/A")

# make chart
Pop_series <- ggplot(popular_df) +
  geom_line(aes(x = CheckoutMonth,
               y = Checkouts,
               group = Series,
               colour = Series),
            show.legend = TRUE) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_fill_brewer(palette= 1) + 
  labs(x = "Checkout Month", 
       y = "Amount of Checkouts",
       title = "Amount of Checkouts Per Month For Each Victoria Schwab Series")

# call chart
Pop_series
```
### Chart Summary
describe why you included the chart, and what patterns emerged

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors ?
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

``````{r, echo = FALSE, warning = FALSE, message = FALSE}
# Question: Is Victoria Schwabs popularity growing or decreasing? If decreasing, which year was her most popular? Which medium is the most popular and how has this changed over the years?

#load Libraries
library("dplyr")
library("ggplot2")

# load Data Frame
Schwab_df <- read.csv("Schwab_Checkouts_by_Title.csv", stringsAsFactors = F)

# modify Data Frame
most_checked_df <- Schwab_df %>%
  filter(CheckoutYear != "2023") %>%
  filter(CheckoutYear != "2011") %>%
  group_by(CheckoutYear, MaterialType) %>%
  summarise(Checkouts = sum(Checkouts))
  
  


# make chart
Checkouts_year <- ggplot(most_checked_df) +
  geom_line(aes(x = CheckoutYear,
                y = Checkouts,
                fill = MaterialType,
                colour = MaterialType),
                show.legend = TRUE) + 
  scale_x_continuous(breaks = seq(2012, 2022, 1)) +
  labs(x = "Year of Checkout", 
       y = "Amount of Checkouts",
       title = "Victoria Schwab's Popularity Over the Years by Medium")

# call chart
Checkouts_year
  
```
### Chart Summary
describe why you included the chart, and what patterns emerged

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

``````{r, echo = FALSE, warning = FALSE, message = FALSE}
# Question: Which Victoria Schwab book is the most popular in the past year?

#load Libraries
library("dplyr")
library("ggplot2")

# load Data Frame
Schwab_df <- read.csv("Schwab_Checkouts_by_Title.csv", stringsAsFactors = F)

Schwab_df$Title <- tolower(Schwab_df$Title)
Schwab_df$Title[str_detect(Schwab_df$Title, "vicious")] <- "Vicious"
Schwab_df$Title[str_detect(Schwab_df$Title, "a conjuring of light")] <- "A Conjuring of Light"
Schwab_df$Title[str_detect(Schwab_df$Title, "a gathering of shadows")] <- "A Gathering of Shadows"
Schwab_df$Title[str_detect(Schwab_df$Title, "new beginnings")] <- "New Beginnings"
Schwab_df$Title[str_detect(Schwab_df$Title, "broken ground")] <- "Broken Ground"
Schwab_df$Title[str_detect(Schwab_df$Title, "this savage song")] <- "This Savage Song"
Schwab_df$Title[str_detect(Schwab_df$Title, "the unbound")] <- "The Unbound"
Schwab_df$Title[str_detect(Schwab_df$Title, "a darker shade of magic")] <- "A Darker Shade of Magic"
Schwab_df$Title[str_detect(Schwab_df$Title, "the near witch")] <- "The Near Witch"
Schwab_df$Title[str_detect(Schwab_df$Title, "second chances")] <- "Second Chances"
Schwab_df$Title[str_detect(Schwab_df$Title, "last wishes")] <- "Last Wishes"
Schwab_df$Title[str_detect(Schwab_df$Title, "the archived")] <- "The Archived"
Schwab_df$Title[str_detect(Schwab_df$Title, "bridge of souls")] <- "Bridge of Souls"
Schwab_df$Title[str_detect(Schwab_df$Title, "city of ghosts")] <- "City of Ghosts"
Schwab_df$Title[str_detect(Schwab_df$Title, "conjuro de luz")] <- "A Conjuring of Light"
Schwab_df$Title[str_detect(Schwab_df$Title, "extraordinary")] <- "ExtraOrdinary"
Schwab_df$Title[str_detect(Schwab_df$Title, "gallant")] <- "Gallant"
Schwab_df$Title[str_detect(Schwab_df$Title, "la vida invisible de addie larue")] <- "The Invisible Life of Addie LaRue"
Schwab_df$Title[str_detect(Schwab_df$Title, "the invisible life of addie larue")] <- "The Invisible Life of Addie LaRue"
Schwab_df$Title[str_detect(Schwab_df$Title, "our dark duet")] <- "Our Dark Duet"
Schwab_df$Title[str_detect(Schwab_df$Title, "tunnel of bones")] <- "Tunnel of Bones"
Schwab_df$Title[str_detect(Schwab_df$Title, "vengeful")] <- "Vengeful"


# modify Data Frame
popular_df <- Schwab_df %>%
  filter(CheckoutYear == "2022") %>%
  group_by(Title) %>%
  summarise(Checkouts = sum(Checkouts))

# make chart
Pop_book <- ggplot(popular_df) +
  geom_col(aes(x = Checkouts,
               y = Title,
               fill = Checkouts)) +
  labs(x = "Amount of Checkouts", 
       y = "Victoria Schwab Books",
       title = "Amount of Checkouts For Each Victoria Schwab Book")

# call chart
Pop_book
```
### Chart Summary
describe why you included the chart, and what patterns emerged